---
title: "Midterm Lab"
author: "STAT 359 Data Visualization"
output: html_document
---

# Overview

For reference, the graphics were generated to be 6 inches by 6 inches. Tasks 1 & 2 ask you to replicate plots and provide some basic insight into what the graphics are communicating. Task 3 asks you to provide short descriptions/explanations regarding issues and concepts in `ggplot2`. 

# Datasets 

The `appendix.csv` file is from Kaggle.com (https://www.kaggle.com/edx/course-study/version/1).

# Tasks

0a. Describe each of the following:

1. `ggplot()`
2. `aes()`
3. `geoms`
4. `stats`
5. `scales`
6. `theme()`

<br>

0b. Explain the difference between using this code `geom_point(aes(color = VARIABLE))` as opposed to using `geom_point(color = VARIABLE)`.

``

<br>

1a. Run the below code and answer the below questions.

```{r, message=FALSE, warning=FALSE}
library(tidyverse)

online_course_data <- read_csv('appendix.csv', na = '---') %>% 
  mutate(Institution = factor(Institution),
         `Honor Code Certificates`  = factor(`Honor Code Certificates`),
         `Course Subject` = factor(`Course Subject`))

dim(online_course_data)
glimpse(online_course_data)
```

<br>

1b. How many courses are there per subject for each instiution? 

Notes:
 * Change the y-label to be more descriptive than "count"

```{r, message=FALSE, warning=FALSE}
ggplot(online_course_data, aes(`Course Subject`, fill = Institution)) +
  geom_bar() +
  labs(y = 'Number of Courses')

# OR

ggplot(online_course_data, aes(`Course Subject`)) +
  geom_bar() +
  facet_wrap(~ Institution) +
  theme(legend.position="bottom") +
  labs(y = 'Number of Courses')

# OR

ggplot(online_course_data, aes(Institution)) +
  geom_bar() +
  facet_wrap(~ `Course Subject`, scales = 'free_y') +
  theme(legend.position="bottom") +
  labs(y = 'Number of Courses')
```

<br>

1c. In one plot, show the `Median Age` of Course Subjects (including the 25th, 50th, and 75th percentiles) broken out by both Year and Institution.

```{r, message=FALSE, warning=FALSE}
ggplot(online_course_data, aes(`Course Subject`, `Median Age`)) +
  geom_boxplot() +
  facet_grid(Year ~ Institution)

# OR

ggplot(online_course_data, aes(`Course Subject`, `Median Age`, fill = Institution)) +
  geom_boxplot() +
  facet_wrap(~ Year)

# OR

ggplot(online_course_data, aes(`Course Subject`, `Median Age`, fill = Year)) +
  geom_boxplot() +
  facet_wrap(~ Institution)
```

<br>

2d. Recreate the below visualization.

```{r, message=FALSE, warning=FALSE}
ggplot(online_course_data, aes(`% Certified`, `Median Hours for Certification`, color = Institution)) +
  geom_smooth(se = FALSE, method = 'lm', show.legend = FALSE) +
  geom_point() +
  scale_y_log10() +
  scale_x_log10() +
  labs(title = 'Hours for Certification versus % Certified')
```

<br>

2e. Create a visualization which plots `Audited (> 50% Course Content Accessed)` against `% Bachelor's Degree or Higher` and displays the name of the course instead of a dot, broken out by the course subject.

```{r, message=FALSE, warning=FALSE}
ggplot(online_course_data, aes(`Audited (> 50% Course Content Accessed)`, `% Bachelor's Degree or Higher`, label = `Course Title`)) +
  geom_text(check_overlap = TRUE, vjust = 'inward', hjust = 'inward') +
  scale_x_log10() +
  facet_wrap(~ `Course Subject`)
```