---
title: "Lab 1"
author: "STAT 359 Data Visualization"
output: html_document
---

# Overview

The intent of this lab is to help students determine how well they actually know R. You should begin reading chapters 1 and 2 in Hadley Wickham's *ggplot2: Elegant Graphics for Data Analysis*.

# Datasets 

We will be using the `catsvdogs.txt`, `titanic.txt`, and `mmDATA.txt` datasets for this lab. The datasets can be found on Canvas in **Files** in either the `datasets_master` directory or `getting_started\datasets` directory. 

# Tasks

Use the `catsvdogs` dataset to complete the following. Should begin by exploring the data and reading *catsvdogs_codebook.txt* file.

If you receive an error saying something along the lines of `Error in library(someLibrary) : there is no package called 'someLibrary'...` then in your console type `install.packages('someLibrary')`.
  
```{r, message=FALSE, warning=FALSE}
# Import libraries
library(tidyverse)
library(ggthemes)
library(knitr)
library(kableExtra)

# Set theme (don't worry about this now... you'll learn about it later)
theme_set(theme_minimal())

# Import data (don't worry about this now... you'll learn about it later)
catsvdogs_data <- read_delim('catsvdogs.txt', '|')

titanic_data <- read_delim('titanic.txt', '|') %>% 
  mutate(class = factor(class),
         adult = factor(adult),
         sex = factor(sex),
         survived = factor(survived))

mm_data <- read_delim('mmDATA.txt', '|')
```

<br>

0a. What do you think you should do when you encounter an error in RStudio? 

`1. Read the error.`
`2. Try to understand the error.`
`3. Copy and paste the error in Google.`

<br>

0b. How do you install new packages/libraries in R? (hint: Google: "how do I install new packages/libraries in R?")

`install.packages('somePackage')`

<br>

0b. What does the `factor` function do above? (hint: Google "factor() in R")

`It "encodes" a vector/column as a categorical variable.`

<br>

### Cats versus Dogs data

1. Construct a stem plot (histogram or density plot) of the percent of households with pets. What do you observe?

```{r, message=FALSE, warning=FALSE}
# Response: 50% of more of households in every state, except one (D.C), have pets
ggplot(catsvdogs_data) +
  geom_histogram(aes(percent_households_with_pets), fill = 'black')
```

<br>

2.  Which of the contiguous states (including DC) has the highest/lowest rate of pet ownership?

```{r, message=FALSE, warning=FALSE}
# Highest rate of pet ownership
(catsvdogs_data %>% 
  mutate('pet_owners' = percent_cat_owners + percent_dog_owners) %>% 
  arrange(pet_owners) %>% 
  select(location, pet_owners))[nrow(catsvdogs_data),] %>% 
  kable() %>% 
  kable_styling(full_width = F, position = "left")

# Lowest rate of pet ownership
(catsvdogs_data %>% 
    mutate('pet_owners' = percent_cat_owners + percent_dog_owners) %>% 
    arrange(pet_owners) %>% 
    select(location, pet_owners))[1,] %>% 
  kable() %>% 
  kable_styling(full_width = F, position = "left")
```

<br>

3. Decide on a process to classify each state and DC as preferring cats or dogs. Implement your process and produce a list of states preferring dogs and a list for those that prefer cats.

```{r, message=FALSE, warning=FALSE}
pet_preference <- catsvdogs_data %>% 
  mutate('pet_preference' = ifelse(percent_cat_owners >= percent_dog_owners, 'cats', 'dogs')) %>% 
  select(location, pet_preference)

# These states prefer dogs:
filter(pet_preference, pet_preference == 'dogs') %>% 
  select(location) %>% 
  kable() %>% 
  kable_styling(full_width = F, position = "left")

# These states prefer cats:
filter(pet_preference, pet_preference == 'cats') %>% 
  select(location) %>% 
  kable() %>% 
  kable_styling(full_width = F, position = "left")
```

<br>

4. As part of your exploration of the data, you should have created tables for each of the variables. If not, do so now. Comment on the tables.

<br>

Response: the table contains binary categorical variables. I've made these variables factors during the import process.

<br>

### Titanic data

<br>

5. Construct a bar graph of the variable class.

```{r, message=FALSE, warning=FALSE}
ggplot(titanic_data) + 
  geom_bar(aes(class), fill = 'black')
```

<br>

6. Build a table of the variable class by the variable survived. What do you observe?

```{r, message=FALSE, warning=FALSE}
(survived_class_table <- table(titanic_data$class, titanic_data$survived))
```

<br>

7. Construct a mosaic plot using the package graphics (function mosaicplot). (hint: the input will be the above table built in question 6)

```{r, message=FALSE, warning=FALSE}
mosaicplot(survived_class_table)
```

The middle class had more survivors proportional to the lowest and highest classes.

<br>

8. Explore any other questions you may have.

```{r, message=FALSE, warning=FALSE}
mosaicplot(table(titanic_data$sex, titanic_data$survived))
```

<br>

### mmDATA data

<br>

9. Construct a bar plot of the colors.

```{r, message=FALSE, warning=FALSE}
mm_data_for_graph <- gather(select(mm_data, -source))

ggplot(mm_data_for_graph) +
  geom_bar(aes(value, fill = key))
```

<br>

10. Plot the proportion of each color across time.

```{r, message=FALSE, warning=FALSE}
colors_over_time <- mm_data %>% 
  mutate('source_updated' = str_sub(source, 3),
         sum = blue + orange + green + yellow + brown + red,
         blue_prop = blue / sum,
         orange_prop = orange / sum,
         green_prop = green / sum, 
         yellow_prop = yellow / sum,
         brown_prop = brown / sum,
         red_prop = red / sum) %>% 
  select(source_updated, blue_prop, orange_prop, green_prop, yellow_prop, brown_prop, red_prop) %>% 
  group_by(source_updated) %>% 
  summarise(blue = mean(blue_prop),
            orange = mean(orange_prop),
            green = mean(green_prop),
            yellow = mean(yellow_prop),
            brown = mean(brown_prop),
            red = mean(red_prop))

colors_over_time %>% gather(key, value, blue, orange, green, yellow, brown, red) %>% 
  ggplot(aes(x = source_updated, y = value, group = key, color = key)) +
  geom_line()
```